---
title: "HW 1 - Data visualization"
author: "INSERT NAME"
editor: visual
---

## Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(openintro)
```

## Exercise 1

```{r}
#| label: price-garage

duke_forest |>
  mutate(garage = if_else(str_detect(parking, "Garage"),   "Garage", "No garage")) # |>
  # ggplot(...)
```

Add narrative here...

{{< pagebreak >}}

## Exercise 2

```{r}
#| label: add-ex2-label-here

# add code here
```

Add narrative here...

{{< pagebreak >}}

## Exercise 3

```{r}
#| label: load-data
#| message: false

brfss <- read_csv("data/brfss.csv")
```

```{r}
#| label: add-ex3-label-here

# add code here
```

Add narrative here...

{{< pagebreak >}}

## Exercise 4

```{r}
#| label: add-ex4-label-here

# add code here
```

Add narrative here...

{{< pagebreak >}}

## Exercise 5

```{r}
#| label: add-ex5-label-here

# add code here
```

Add narrative here...

{{< pagebreak >}}

## Exercise 6

(a) 

    -   The gg in the name of the package ggplot2 stands for \_\_\_.
    -   If you map the same continuous variable to both `x` and `y` aesthetics in a scatterplot, you get a straight \_\_\_ line. (Choose between "vertical", "horizontal", or "diagonal".)

(b) Add narrative here...

```{r}
#| label: style

ggplot(data=mpg,mapping=aes(x=drv,fill=class))+geom_bar() +scale_fill_viridis_d()
```

(c) Add narrative here...
